The evolution of the transcriptome of mammalian placentas

### 2016-11-11 manuscript/manuscript.Rnw
start discussing morphology of placenta attachment et al

The core placenta transcriptome is significantly enriched in genes
which are involved in multiple biological pathways, including multiple
pathways related to translation (ribosome, translation, signal
recognition particle targeting to the rough endoplasamic reticulum,
etc.) (Supplemental \cref{tab:core_placenta_pathway_analysis}). Beyond
translation, the next most abundant set of significantly enriched
molecular pathways are related to the migration of the trophoblast
into the endometrium, including the EGFR1 pathway and the Actin Y
pathway. Two other pathways, the mPR/Progesterone pathway and Rho cell
motility pathways, feed into the Actin Y pathway as well. Supplemental
\cref{fig:egf_kegg_pathway} shows the EGF pathway in humans with the
median expression in FPKM of genes in the pathway shown. Notably, many
of the genes on the GRB2->MYC pathway are strongly expressed; their
activation leads to an increase in migration and invasion in response
to EGF binding to EGFR1. The Actin cytoskeleton pathway is depicted in
Supplemental \cref{fig:kegg_actin}, which shows that \textit{FN1} and
\textit{ITG} are highly expressed, which activate Rac, and potentially
lead to the formation of focal adhesions and lamellipodia formation
through their inhibition of GSN and HSPC300.

@@

@@

@@

>>>>

The core placenta transcriptome is significantly enriched in genes
which are involved in multiple biological pathways, including multiple
pathways related to translation (ribosome, translation, signal
recognition particle targeting to the rough endoplasamic reticulum,
etc.) (Supplemental \cref{tab:core_placenta_pathway_analysis}). Beyond
translation, the next most abundant set of significantly enriched
molecular pathways are related to the migration of the trophoblast
into the endometrium, including the EGFR1 pathway and the Actin Y
pathway. Two other pathways, the mPR/Progesterone pathway and Rho cell
motility pathways, feed into the Actin Y pathway as well. Supplemental
\cref{fig:egf_kegg_pathway} shows the EGF pathway in humans with the
median expression in FPKM of genes in the pathway shown. Notably, many
of the genes on the GRB2→MYC pathway are strongly expressed; their
activation leads to an increase in migration and invasion in response
to EGF binding to EGFR1. The Actin cytoskeleton pathway is depicted in
Supplemental \cref{fig:kegg_actin}, which shows that \textit{FN1} and
\textit{ITG} are highly expressed, which activate Rac, and potentially
lead to the formation of focal adhesions and lamellipodia formation
through their inhibition of GSN and HSPC300.

@@

We correlated different aspects of placenta morphology, including
shape, intimacy, and interdigitation with the mRNA transcription
profiles of all 1:1 human orthologs with \ac{FPKM} greater than 10 in
one species, and \ac{FPKM} greater than 1 at least two species.

@@

Supplemental \cref{tab:shape_genes} shows the top 50 genes which are
associated with the shape of placenta attachment, as well as the shape
in which the difference is most significant.

\textit{EREG} -- ligand of EGFR, known to be expressed in some
placentas but apparently not present in the term placentas of \human.

\textit{UNC} -- totally unknown

@@

Genes in multiple immune system-related pathways are altered in increasing degrees of intimacy
